\chapter{Introduction}

\section{Motivation}

Our world is becoming more interconnected than ever before, with over 10 billion devices connected to WLAN (Wireless Local Area Network) in 2017 and projections placing this number around 20 billion by 2021\cite{devices-on-wlan-2017}. The further introduction of the Internet of Things (IoT) and many other smart devices such as home assistants, like Google Home\footnote{Google Home <\url{https://store.google.com/product/google_home/}>} or Amazon Alexa\footnote{Amazon Alexa <\url{https://www.amazon.com/b/ref=gbpp_itr_m-2_2551_16067214?node=16067214011&ie=UTF8}>}, along with smart home devices ranging from thermostats and light bulbs to refrigerators and washing machines put these predictions at over 35 billion devices connected to the internet by 2021\cite{iot-devices}.

For the majority of people, being connected to the internet at every point in their day, directly or indirectly, has become a matter of fact. These developments have made the need for reliable and fast transmission of data more relevant than ever before.

The problem of real-time communication is increasingly important as bandwidth and connections speeds increase, allowing for more interactivity between users and computer systems. More and more systems require increasing amounts of communication in this age of an ever increasing number of web-based applications taking the place of traditional desktop ones. With services feeding customized content to a large number of different users in a more interactive and immediate manner, traditional methods prove too slow and cumbersome.

\section{Goals}

The goals of this thesis are to create a system that can serve as a framework and foundation to be built upon and expanded for usage in modern applications that need fast communication among potentially large numbers of devices, and a simple sample application in order to test the system's functionality.

\section{Usage Scenarios}

The system designed and implemented as part of this thesis is meant to server as an extensible framework for modern applications, as mentioned above. Some of the many possible usages for the system can be instant messaging (IM), media (video or audio) streaming, communication within IoT systems, smart devices and more.
